A global compilation reveals divergent biomass partitioning between plant functional types

### 2014-10-23 manuscript/draftpaper.Rnw

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
%\usepackage{lineno}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\bibliographystyle{plain}

@@

%\linenumbers

\title{Specific leaf mass explains difference in biomass partitioning between woody plant functional types}

@@

\textbf{The partitioning of biomass in terrestrial vegetation between leaves, stems and roots is a critical component of dynamic global vegetation models (DGVMs). Biomass partioning is consistently shown to be one of the most uncertain and important drivers for the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, and global predictions of carbon storage \cite{ise_comparison_2010}. Little is known about direct and indirect effects of climate on biomass partitioning, largely due to the lack of a global individual-based biomass dataset. Here, we analyze the largest database on plant biomass to date. We show that partitioning of biomass to leaves vs. woody stems is strongly dependent on the plant functional type. Evergreen gymnosperms maintain three times more leaf biomass at a given total aboveground biomass than deciduous angiosperms, with evergreen angiosperms taking an intermediate value. We found that specific leaf mass, an indicator of leaf construction cost, explains these differences so that the PFTs converged in the ratio  of leaf area to aboveground biomass. We found weak effects of mean annual temperature and no effects of precipitation on biomass partitioning, suggesting that climate-related differences are captured largely by a shift in PFT. These results point to a simple algorithm for biomass partitioning in the world’s forests.}

>>>>

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{lineno}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\bibliographystyle{plain}

@@

\linenumbers

\title{A global compilation reveals divergent biomass partitioning between plant functional types }

@@

\textbf{The partitioning of biomass in terrestrial vegetation between leaves, stems and roots is a critical component of dynamic global vegetation models (DGVMs). Biomass partioning is consistently shown to be one of the most uncertain and important drivers for global predictions of carbon storage \cite{ise_comparison_2010}, the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, and the CO$_2$ response of net primary production \cite{dekauwe2014does} . Little is known about direct and indirect effects of climate on biomass partitioning, largely due to the lack of a global individual-based biomass dataset. Here, we analyze the largest database on plant biomass to date [cite]. We show that partitioning of biomass to leaves vs. woody stems is strongly dependent on the plant functional type. Evergreen gymnosperms maintain three times more leaf biomass at a given total aboveground biomass than deciduous angiosperms, with evergreen angiosperms taking an intermediate value. We found that specific leaf mass, an indicator of leaf construction cost, explains these differences so that the PFTs converged in the ratio  of leaf area to aboveground biomass. We found weak effects of mean annual temperature and no effects of precipitation on biomass partitioning, suggesting that climate-related differences are captured largely by a shift in PFT. These results point to a simple algorithm for biomass partitioning in the world’s forests.}
